paper1_figures

### 2023-09-18 paper1_markdown.Rmd
18.09 version

@@

Reaching a better understanding of mechanism underlying cooperation in social dilemmas is even more important when considering that many everyday situations have a social dilemma structure. An example is the climate crisis that we are actually facing: engaging in pro-environmental behaviours to reduce our ecological footprint is often more complicated and less advantageous than acting without worrying about environment [@Griskevicius2010]. However, if everyone defects by choosing the easy way, consequences will be (and already are) catastrophic. Another (less tragic) example is provided by @Kollock1998 and involves public television: the best outcome at the individual level would be to have access to the programs without paying the fee, but if nobody pays it, then public television will stop existing. + @Kollock1998: page 183, many of the problems that we face are, at their core, social dilemmas. A third example of real-life social dilemmas are commons.

@@

To conduct this scoping review, different papers published in academic journals were analyzed in response to the research questions. To gather the papers, a Boolean search methodology was used in different scientific search engines such as Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", "commons", "cooperation", "conflicts", "reciprocity", "trust" **, "intergroup"**, and they were paired with "AND" and "OR" Boolean operators. The research was made during the month of August, 2023. According with @Munn2022, the type of source was not a selection criteria, therefore experimental and qualitative studies, as well as meta-analysis and systematic reviews are included. This allows us to have an idea as complete as possible about the existing literature on cooperation in commons and social dilemmas. Moreover, there were not chronological or geographical limitations, once again to provide a broader vision on the topic. There were however some selection criteria concerning the papers' content. First of all, as mentioned in the introduction, this review will focus on individual and group dynamic that can enhance or decrease cooperation in a social dilemma situation (i. e. commons). For this reason, all the papers investigating how an external entity/institution can shape cooperation were not included. Secondly, papers that studied social dilemmas with an agent-based model methodology were also excluded. **include explication**

@@

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group (e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020). For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, the link between group identification and cooperation is the same for individualist and collectivist people [@Chen2007], which is a further confirmation of the strength of this effect. Finally, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015; @Aksoy2019].

@@

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In a social dilemma experiment, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment using a public resource game, intergroup competition reduced resources consumption (and therefore enhanced cooperation), while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition per se.

@@

>>>>

introduction of the societal problem = cooperation (= renonce to the individual most advantageous choice for the common good) is the key to resolve many societal problems (e.g. covid vaccination, climate change)

@@

Reaching a better understanding of mechanism underlying cooperation in social dilemmas is even more important when considering that many everyday situations have a social dilemma structure. An example is the climate crisis that we are actually facing: engaging in pro-environmental behaviours to reduce our ecological footprint is often more complicated and less advantageous than acting without worrying about environment [@Griskevicius2010]. However, if everyone defects by choosing the easy way, consequences will be (and already are) catastrophic. Another (less tragic) example is provided by @Kollock1998 and involves public television: the best outcome at the individual level would be to have access to the programs without paying the fee, but if nobody pays it, then public television will stop existing. + @Kollock1998: page 183, many of the problems that we face are, at their core, social dilemmas. A third example of real-life social dilemmas are commons. (**USE THIS AT THE VERY BEGINNING TO INTRODUCE THE SOCIETAL PROBLEM?**)

@@

To conduct this scoping review, different papers published in academic journals were analyzed in response to the research questions. To gather the papers, a Boolean search methodology was used in different scientific search engines such as Scopus, Web of Science and ScienceDirect. The key phrases were "social dilemma", "public good", "common resource", "commons", "cooperation", "conflicts", "reciprocity", "trust" **, "intergroup"**, and they were paired with "AND" and "OR" Boolean operators. The research was made during the month of August, 2023. According with @Munn2022, the type of source was not a selection criteria, therefore experimental and qualitative studies, as well as meta-analysis and systematic reviews are included. This allows us to have an idea as complete as possible about the existing literature on cooperation in commons and social dilemmas. Moreover, there were not chronological or geographical limitations, once again to provide a broader vision of the topic. There were however some selection criteria concerning the papers' content. First of all, as mentioned in the introduction, this review will focus on individual and group dynamic that can enhance or decrease cooperation in a social dilemma situation (i. e. commons). For this reason, all the papers investigating how an external entity/institution can shape cooperation were not included. Secondly, papers that studied social dilemmas with an agent-based model methodology were also excluded. **include explication**

@@

Social identity have a consensual positive effect on cooperation: the more individuals feel a strong identification with a group, the more they cooperate with other members of the group [e.g. @Arora2016; @Noonan2016; @Gomez-Ruiz2020]. For example, in a field study about community-organized dog parks in the United States, @Noonan2016 found that a strong sense of community was related to a higher amount of voluntary activities to maintain the parks in good conditions. This is explained by the fact that, when people strongly identify with a social group, they are more concerned about their group's well-being and reputation [@VanVugt2009] and they accord more importance to group-related goal [@Arora2016]. A strong social identity also maintains a high level of cooperation over time [@Zhang2019]. Moreover, the link between group identification and cooperation is the same for individualist and collectivist people [@Chen2007], which is a further confirmation of the strength of this effect. Finally, independently of the strength of group identification, people cooperate more with other people in the same group than with out-group people [@Dorrough2015; @Aksoy2019].

@@

Another well replicated result is the positive effect of intergroup competition on intragroup cooperation. In fact, in-group identification can be enhanced via a competition between different groups, because in such a context individual's and group's interests are aligned [@Puurtinen2009]. In a social dilemma experiment, these authors demonstrated that cooperation was higher in an intergroup-competition condition than in a non-competitive condition. Moreover, when participants had to compete against other groups, they reported higher levels of anger toward in-group free-riders and they felt more guilty after a defection decision. In another experiment using a public resource game, intergroup competition reduced resources consumption, while individual, dispositional competitiveness predicted an higher level of harvesting [@Nockur2020]. This shows that the increase of cooperation via competition is due to group membership factors, and not to competition per se.

@@

## Social status???

retrieve work on gender that says that gender effects are probably in part impacted by status?

# Group factors

## Reward and punishment

Possible introduction: Incentives are divided in reward and punishments. Punishment has received a greater attention than rewards [@Balliet2011a; @VanDijk2016]. The positive effect of punishment in promoting cooperation is well confirmed ...

On the other side, results about rewarding are not so consensual ...

@Chen2022: reward, but not punishment, increased cooperation in the following round for both high and low benefit members

@Nockur2021: punishment (three different types) increase cooperation + the presence of a punishment system increased the perception of fairness + satisfaction was higher in peer and democratic punishment than in centralized and no punishment condition

@Kingsley2016: the possibility of punish others increased the cooperation in homogeneous groups (= where all members started with the same endowment), but not in heterogeneous groups

@Milinski2012: punishment is worthy on the long run (straight after, the ones who punished are the ones with the lowest income) + reward is less efficient than punishment (punish: at the beginning, to establish cooperation, reward is a cost all along because once cooperation is established you always have to continue rewarding it) + punishment can however entail a second-order social dilemma (defection = do not punish but having advantages for the fact that everybody cooperates thanks to punishers)

@Xiao2016: higher cooperation if only the cooperator can punish the defector (vs everybody can punish), and in particular for deterministic (vs stochastic) game\
when both can punish, anti-social punishment is present --\> effect = a person who cooperated and received a punishment in t is less likely to cooperate in t+1 (proportional)\
in stochastic games, the occurrence of a negative event is not a predictor for punishing (= based on other's defection, independently of the outcome)\
punishment of non-investor is smaller in stochastic (vs deterministic) games

@Balliet2011a: PUN: positive effect on cooperation in social dilemmas\
stronger if the punishment was costly (vs free)\
stronger if decentralized (vs centralized)\
stronger if SD with partners (vs strangers)\
stronger if iterated (vs one-shot)\
REWARD: positive effect on cooperation in social dilemmas\
stronger if costly vs free\
no effect of de-centralized\
stronger if iterated (vs one-shot)\
stronger in the resource dilemma (vs public good or prisoner dilemma)\
OVERALL: no difference on the strength of these effects

@Jacquet2011: the expectation of honour and shame enhance cooperation

@Rockenbach2019: successful groups punish, but with caution and not in early stages (see Excel for more results)

@VanDijk2016: results are both in the introduction and result case

exclusion as a form of punishment

## Communication

## Group size

## Social Norms
